\documentclass[11pt,a4paper,roman]{moderncv}      
\usepackage[english]{babel}

\moderncvstyle{classic}                            
\moderncvcolor{black}                            

% character encoding
\usepackage[utf8]{inputenc}

% adjust the page margins
\usepackage[scale=0.80]{geometry}

% personal data
\name{Swapna Roy}{}
\email{royswapna022@gmail.com}
\phone[mobile]{+91 8334826644}


\begin{document}

\recipient{To}{The Counsellor, \\
               10th Ward,\\ 
               Noapara Kalibari Road, Barasat, 700124}
\date{April, 25 2023}

\opening{\textbf{Sub: Request for closure of open-drain running along the entrance of house.}}
\closing{Your Sincerly, \vspace{-1em}}


% \enclosure[Enclosed (Documents/Certificates)]{
%               \\ 1. Appointment fixed to CMC Vellore.\\
%                  2. Kolkata - Bangalore Flight details.
%                  }

\makelettertitle



Respected Sir,
\\
%references such as what and how you got this information
\vspace{1em}
I am writing to you with utmost urgency and concern regarding the open drain structure running directly along the entrance of our house.
This structure is being used by the \textbf{Originator} of the structure (Mr. Bagchi and his families) 
to transfer human excretion, and the stench that emanates from it is unbearable.
The air we breathe is no longer safe, and we are living in a state of constant discomfort and distress.


%narrate why you want to apply
\vspace{1em}
This open drain structure is not only a health hazard, but it is also a 
violation of our basic rights to live in a clean and hygienic environment. 
The fact that it is located just outside our home entrance area and along a Shiv Temple
makes it even more concerning, as it poses a serious threat to the health and well-being of our families.

\vspace{1em}
I urge you to take immediate action to close or demolish this open drain structure.
The situation is dire, and we cannot continue to live in these conditions. I implore you to use your authority to enforce the necessary measures to ensure that this issue is resolved as soon as possible.


% %conclusion
\vspace{1em}
We are counting on you to take swift and decisive action to address this matter. Our health and safety depend on it.
Thank you for your understanding.
\vspace{0.5cm}


\makeletterclosing

Cosigned by \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\

\vspace{1em}
Signed by \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\

\end{document}
